Clustering time series based on probability distributions across temporal granularities

### 2021-11-18 main.Rmd
say what the number in Figure 7 represents

Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter. The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual. We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation. Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates. Additionally, there were missing values in many series. (The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.

Huge data sets present more complications for clustering. Clustering algorithms work well when there are well-separated clusters, with no nuisance variables or nuisance observations. When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables. Some customers may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters. For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply. Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering. The customers are displayed in two columns of $12$ for space reasons. Each row, of each column, represents the profile of a single customer across different variables. This is often a good approach to tackling a big analysis task, to start with a simpler task. The approach, however, is applicable to all customers.

@@

```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot. Group P-1 and P-2 correspond to Q-1 and Q-2 respectively. Cluster P-3 splits into Q-3, Q-4 and Q-5. P-1 (customers 16, 18) is characterized by high values of $wpd$ on $wnwd$. P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$. P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$. For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance of $moy$ and $wnwd$.", fig.height=6}
```

>>>>

Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter. The full data contains half-hourly general supply in kWh for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual. We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation. Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates. Additionally, there were missing values in many series. (The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.

Huge data sets present more complications for clustering. Clustering algorithms work well when there are well-separated clusters, with no nuisance variables or nuisance observations. When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables. Some customers may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters. For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply. Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering. The customers are displayed in two columns of $12$ for space reasons. Each row, of each column, represents the profile of a single customer across different variables. Each customer is associated with an identifier of the form [$a$-$b$], where $a \in \{1, 2, \dots, 24\}$ represents the customer-prototype id and $b \in \{1, 2, \dots, 5\}$ indicates the label of the prototype in which a customer was placed. This is often a good approach to tackling a big analysis task, to start with a simpler task. The approach, however, is applicable to all customers.

@@

```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD approach showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot. P-1 and P-2 groups correspond to Q-1 and Q-2, respectively. Cluster P-3 is subdivided into Q-3, Q-4, and Q-5. P-1 (customers 16, 18) is distinguished by high $wpd$ on $wnwd$ values. P-2 has lower $wpd$ than $moy$ and $wnwd$ for $hod$. P-3 operates in the opposite way as P-2, with larger $wpd$ for $hod$ in comparison to $moy$ and $wnwd$. For the $5$ cluster solution, this group is divided into Q3-Q5, which are distinct due to their different relative significance of $moy$ and $wnwd$.", fig.height=6}
```
